Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-22 doc/manuscript/ESSDmanuscript.Rmd
move abstract out of front matter
this is so that statistics can be automatically inserted

affiliation:
  - code: 1
    address: Center for Conservation Ecology, Smithsonian's National Zoo & Conservation Biology Institute, Front Royal, VA, United States
  - code: 2
    address: Forest Global Earth Observatory, Smithsonian Tropical Research Institute, Panama, Republic of Panama
  - code: 3
    address: School of Geography, University of Leeds, Leeds, UK
  - code: 4
    address: Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, United States
  - code: 5
    address: The Nature Conservancy; Arlington VA 22203, USA
  - code: 6
    address: CIRAD, Montpellier, France
abstract: |
  Forests are critical for climate change mitigation and consitute a substantial portion of planned emissions reductions under the 2015 Paris Agreement.
  Yet, the efficacy of greenhouse gas mitigation planning and reporting is dependent upon the quality of available emission factors data, including forest carbon (C) stocks and changes therein.
  Tens of thousands of relevant forest C estimates have been published, yet are not readily accesible to the practitioners compiling national greenhouse gas inventories. 
  Many of these data have, however, been compiled in the Global Forest C database (ForC; https://forc-db.github.io/) and stand to be of value to greenhouse gas accounting if made available through the Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC).
  Here, we develop and document a process for semi-automated transfer of data from ForC into the EFDB, assess the data available and transferred to date, and provide recommendations for improving forest data collection, analysis, and reporting to improve accounting of forest-sector greenhoouse gas emissions and removals.
  We begin by reconciling terminology and mapping ForC fields into EFDB.
  This process required some updates to the ForC database structure, leading to the release of a new version of ForC (v4.0; described here).
  At the time of writing, ForC contained ## values that would qualify for inclusion in the EFDB, ## of which have been transferred to date.
  (Some analysis of representation/ gaps.)
  <!-- -->
  In the future, forest C estimates in EFDB can be improved through targetted research to fill critical gaps, reporting of information required by IPCC, and continued submission of data from scientific publications to the EFDB.

@@

>>>>

affiliation:
  - code: 1
    address: Center for Conservation Ecology, Smithsonian's National Zoo & Conservation Biology Institute, Front Royal, VA, United States
  - code: 2
    address: Forest Global Earth Observatory, Smithsonian Tropical Research Institute, Panama, Republic of Panama
  - code: 3
    address: School of Geography, University of Leeds, Leeds, UK
  - code: 4
    address: Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, United States
  - code: 5
    address: The Nature Conservancy; Arlington VA 22203, USA
  - code: 6
    address: CIRAD, Montpellier, France
abstract: |

@@

**Abstract.**
Forests are critical for climate change mitigation and consitute a substantial portion of planned emissions reductions under the 2015 Paris Agreement.
Yet, the efficacy of greenhouse gas mitigation planning and reporting is dependent upon the quality of available emission factors data, including forest carbon (C) stocks and changes therein.
Tens of thousands of relevant forest C estimates have been published, yet are not readily accesible to the practitioners compiling national greenhouse gas inventories. 
Many of these data have, however, been compiled in the Global Forest C database (ForC; https://forc-db.github.io/) and stand to be of value to greenhouse gas accounting if made available through the Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC).
Here, we develop and document a process for semi-automated transfer of data from ForC into the EFDB, assess the data available and transferred to date, and provide recommendations for improving forest data collection, analysis, and reporting to improve accounting of forest-sector greenhoouse gas emissions and removals.
We begin by reconciling terminology and mapping ForC fields into EFDB.
This process required some updates to the ForC database structure, leading to the release of a new version of ForC (v4.0; described here).
At the time of writing, ForC contained ## values that would qualify for inclusion in the EFDB, ## of which have been transferred to date.
(Some analysis of representation/ gaps.)
<!-- -->
In the future, forest C estimates in EFDB can be improved through targetted research to fill critical gaps, reporting of information required by IPCC, and continued submission of data from scientific publications to the EFDB.

### 2022-04-22 doc/manuscript/ESSDmanuscript.tex
move abstract out of front matter
this is so that statistics can be automatically inserted

\begin{abstract}
Forests are critical for climate change mitigation and consitute a
substantial portion of planned emissions reductions under the 2015 Paris
Agreement. Yet, the efficacy of greenhouse gas mitigation planning and
reporting is dependent upon the quality of available emission factors
data, including forest carbon (C) stocks and changes therein. Tens of
thousands of relevant forest C estimates have been published, yet are
not readily accesible to the practitioners compiling national greenhouse
gas inventories. Many of these data have, however, been compiled in the
Global Forest C database (ForC; https://forc-db.github.io/) and stand to
be of value to greenhouse gas accounting if made available through the
Emission Factor Database (EFDB) of the International Panel on Climate
Change (IPCC). Here, we develop and document a process for
semi-automated transfer of data from ForC into the EFDB, assess the data
available and transferred to date, and provide recommendations for
improving forest data collection, analysis, and reporting to improve
accounting of forest-sector greenhoouse gas emissions and removals. We
begin by reconciling terminology and mapping ForC fields into EFDB. This
process required some updates to the ForC database structure, leading to
the release of a new version of ForC (v4.0; described here). At the time
of writing, ForC contained \#\# values that would qualify for inclusion
in the EFDB, \#\# of which have been transferred to date. (Some analysis
of representation/ gaps.) In the future, forest C estimates in EFDB can
be improved through targetted research to fill critical gaps, reporting
of information required by IPCC, and continued submission of data from
scientific publications to the EFDB.
\end{abstract}

>>>>

\textbf{Abstract.} Forests are critical for climate change mitigation
and consitute a substantial portion of planned emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accesible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
accounting if made available through the Emission Factor Database (EFDB)
of the International Panel on Climate Change (IPCC). Here, we develop
and document a process for semi-automated transfer of data from ForC
into the EFDB, assess the data available and transferred to date, and
provide recommendations for improving forest data collection, analysis,
and reporting to improve accounting of forest-sector greenhoouse gas
emissions and removals. We begin by reconciling terminology and mapping
ForC fields into EFDB. This process required some updates to the ForC
database structure, leading to the release of a new version of ForC
(v4.0; described here). At the time of writing, ForC contained \#\#
values that would qualify for inclusion in the EFDB, \#\# of which have
been transferred to date. (Some analysis of representation/ gaps.) In
the future, forest C estimates in EFDB can be improved through targetted
research to fill critical gaps, reporting of information required by
IPCC, and continued submission of data from scientific publications to
the EFDB.
